\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-20 forwards-and-backwards.tex
.

\title{\large{\bf
Looking forwards and backwards through locally regulated populations
}}

@@

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of mature individuals
in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\[
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\]
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation defines $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as discussed later,
differs because offsprings' establishment probability is measured at their new location
rather than their parents', and so differs from $r(x)$.
For the most part, we work with $F$ instead of $\mu$.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        % &=
        % \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align}
\end{definition}

@@

Specialising the results of 
\citet{KURTZ1999103} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.

We write 
$r(x,\eta) \gamma(x,\eta) C(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2C(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

\paragraph{Two-step convergence to PDE:}
Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
we can choose our kernels $\rho_\gamma$, $\rho_r$, and $\rho_F$
in such a way that they depend only on the population in a region close to $x$,
and so we expect that under rather general conditions
solutions of the nonlocal PDE will be close to the corresponding 
classical PDE.
The following propositions provide two concrete situations in which this is true.
In the first, the PDE is a reaction diffusion equation, and we borrow
an idea from~\cite{penington:2017} to express the solutions to both the nonlocal
equation and the classical PDE through a Feynman-Kac formula. 
\begin{proposition}
    \label{prop:nonlocal_to_local}
Recall the notation $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$b(x)$ and $C(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty)>0$,
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}
\begin{remark}
Note that Theorem~\ref{thm:kurtzxiong} guarantees uniqueness of solutions
to equation~(\ref{nonlocalPDEv1}). 
\end{remark}
Our second example is a nonlocal version of a porous medium equation with
logistic growth. That is, 
we consider non-negative solutions to the equation
\begin{equation}
\label{mollified equation}
\partial_t \psi^\epsilon=
\Delta\left(\psi^\epsilon \, \rho^{\epsilon}_\gamma*\psi^\epsilon\right)
+\psi^\epsilon\left(1-\rho^{\epsilon}_\gamma*\psi^\epsilon\right).
\end{equation}
The case without the reaction term (and with $\IR^d$ replaced by a torus) is considered 
by~\cite{lions/mas-gallic:2001} who use
it as a basis for a particle method for numerical solution of the porous medium equation.
Of course this does not quite fit into our framework, since in the notation of our 
population models this would necessitate $\gamma(x,m)=\smooth{\epsilon}m$ which is
not bounded. However, this can be overcome by an additional layer of approximation
(c.f.~our numerical experiments of Section~\ref{beyond linear diffusion})
and we do not allow this to detain us here. Existence and uniqueness of solutions 
to~(\ref{mollified equation}) can be obtained using the approach
of~\cite{lions/mas-gallic:2001}.

@@

\begin{corollary} \label{cor:wavefront}
    In addition to the assumptions of Corollary~\ref{cor:lineages_simple},
    suppose that the population process is described by a travelling wave with 
velocity $\wavespeed$,
    i.e., the population has density
    $\varphi(t, x) = w(x - t \wavespeed)$
    where $w$ solves
    \begin{align*}
        r \DG^* (\gamma w) + w F + \wavespeed \cdot \grad w = 
r\sigma^2\Delta (\gamma w) -\meanq \cdot\grad (\gamma w) +\wavespeed \cdot \grad w =0 .
    \end{align*}
    Then the semigroup $Q_s$
    of the motion of a lineages in this population 
    is time-homogeneous with generator
    \begin{align}
\label{Lgen}
        \Lgen f
        &=
        \sigma^2 r \gamma
        \left(
            \Delta f
            +
            2 \grad \log (\gamma w)
            \cdot \grad f
        \right)
        + (\wavespeed - r\gamma \meanq) \cdot \grad f .
    \end{align}
\end{corollary}

@@

Comparing %\eqref{PME} 
with \eqref{eqn:nonlocal_pde},
we see that to set up a limit in which the population density $\varphi$ follows the 
porous medium equation with logistic growth of~(\ref{PME}),
we need $r=1$,
$\gamma = \varphi$, and $F = 1 - \varphi$.
Consulting equation~\eqref{eqn:mu_defn},
this implies that $\mu_\theta = \max\big(0, (1 + 1/\theta) \varphi - 1/\theta\big)$.
In other words,
establishment is certain
and birth rates increase linearly with population density,
but to compensate, death rates increase slightly faster (also linearly),
as shown in Figure~\ref{fig:pme_waves}.
Alert readers will notice that
the condition from Assumptions~\ref{def:model_setup} 
that $\gamma(x,m)$ be uniformly bounded 
is violated.
This can be corrected by use of a cut-off, and in fact the downwards drift
provided by the logistic control of the population size prevents $m$ from 
getting too big. 
In practice in our simulations we
take discrete time steps of length $dt$ (with $dt$ suitably small),
and have each individual with local density $m$ reproduce and die, respectively, with probabilities
\begin{align*}
    p_\text{birth}(m) = \left(1 - e^{- m dt}\right)
    \qquad 
    p_\text{death}(m) = \left(1 - e^{- (m (1+1/\theta) - 1/\theta) dt} \right) ,
\end{align*}
so that $\gamma(x, m) = p_\text{birth}(m)/dt \approx m$
and
\begin{align*}
    F(x, m) = \theta(\gamma(x, m) - \mu(x, m)) / dt \approx 1 - m .
\end{align*}
Birth and death rates are equal at density $m = 1$,
corresponding to an unscaled density of $N$ individuals per unit area.

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_1234.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_1234.steps}
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta$ small on the left; large on the right.
        Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        (implemented by choosing a small $dt$ taking reproduction probability per time step
        to be $1 - \exp(- dt m)$
        and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        values of $\theta$ are 1 and 100, respectively.
        \label{fig:pme_waves}
    }
\end{figure}

In one dimension, equation~(\ref{PME}) has an explicit travelling wave solution
\begin{align} \label{eqn:pme_wave}
    w^P(t, x)
    :=
    \left( 1 - e^{ \frac{1}{2} (x - x_0 - t) } \right)_+ .
\end{align}
Notice that the wave profile has a sharp boundary at $x = x_0 + t$.
There are also travelling wave solutions with $c>1$ \citep{gilding/kersner:2005},
which lack this property.
However, for initial conditions that decay sufficiently rapidly at infinity,
such as one might use in modelling a population invading new territory,
the solution converges to \eqref{eqn:pme_wave} \citep{kamin/rosenau:2004}.
In Figure~\ref{fig:pme_waves} we show simulations of the individual based model described above.
They display travelling wave solutions qualitatively similar to solutions of~(\ref{PME}),
with better agreement for larger $\theta$ and $N$.

@@

\paragraph{Porous Medium equation with logistic growth:}
Finally, consider equation~(\ref{PME}). Setting $x_0=0$ (for definiteness) and substituting
the form of $w^P$ from equation~\eqref{eqn:pme_wave}
into Corollary~\ref{cor:wavefront},
with $c=1$,
$\gamma(x, m) = m$,
$r(x,w) = 1$,
and $F(x, m) = (1 - m)$,
the generator of the diffusion governing the position of the ancestral lineage 
is, for $x < 0$,
%with $f_x$ denoting the spatial derivative of $f$,
\begin{align*}
    \Lgen_3 f
    &=
        w^P
        \left(
        \partial_{xx} f
         +
         2 \frac{\partial_x((w^P)^2)}{(w^P)^2} \partial_xf
        \right)
        + \partial_xf \\
    &=
        \left(1 - e^{\frac{1}{2} x} \right)
        \partial_{xx}f
        -
        2 e^{\frac{1}{2} x} \partial_xf
        +
        \partial_xf .
\end{align*}
The speed measure corresponding to this diffusion has density
%again for $\xi < 0$,
\begin{align*}
    m(\xi)
    &\propto
        \frac{ 1 }{ 2 (1 - e^{\xi/2}) }
        \exp\left(
            \int_\eta^\xi \left\{
                1 - \frac{e^{x/2}}{1 - e^{x/2}}
            \right\} dx
        \right) \\
    &\propto
        e^\xi\left(1-e^{\xi/2}\right),
        \quad
        \text{for } \xi < 0 ,
\end{align*}
which is integrable and so when suitably normalised gives the unique stationary distribution.
Notice that even though we have the same reaction term as in the Fisher-KPP equation, with this 
form of nonlinear diffusion, at stationarity
the lineage will typically be significantly behind the front, 
suggesting a different genealogy.

>>>>

\title{\large{\bf
Looking forwards and backwards: dynamics and genealogies of locally regulated populations
}}

@@

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of mature individuals
in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\begin{equation} \label{eqn:total_repro}
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\end{equation}
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation defines $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as can be seen from \eqref{eqn:total_repro},
differs because an offspring's establishment probability is measured at their new location
rather than that of their parent.
For the most part, we work with $F$ instead of $\mu$.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f \in C_b^\infty(\IR^d)$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        % &=
        % \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align}
\end{definition}

@@

Specialising the results of 
\citet{KURTZ1999103} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.
We write 
$r(x,\eta) \gamma(x,\eta) \covq(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2\covq(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

@@

\paragraph{Two-step convergence to PDE:}
Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
we can choose our kernels $\rho_\gamma$, $\rho_r$, and $\rho_F$
in such a way that they depend only on the population in a region close to $x$,
and so we expect that under rather general conditions
solutions of the nonlocal PDE will be close to the corresponding 
classical PDE.
The following propositions provide two concrete situations in which this is true.
In the first, the PDE is a reaction diffusion equation, and we borrow
an idea from~\cite{penington:2017} to express the solutions to both the nonlocal
equation and the classical PDE through a Feynman-Kac formula. 
\begin{proposition}
    \label{prop:nonlocal_to_local}
Recall the notation $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$\meanq(x)$ and $\covq(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty)>0$,
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}
\begin{remark}
Note that Theorem~\ref{thm:kurtzxiong} guarantees uniqueness of solutions
to equation~(\ref{nonlocalPDEv1}). 
\end{remark}
Our second example is a nonlocal version of a porous medium equation with
logistic growth. That is, 
we consider non-negative solutions to the equation
\begin{equation}
\label{mollified equation}
\partial_t \psi^\epsilon=
\Delta\left(\psi^\epsilon \, \rho^{\epsilon}_\gamma*\psi^\epsilon\right)
+\psi^\epsilon\left(1-\rho^{\epsilon}_\gamma*\psi^\epsilon\right).
\end{equation}
The case without the reaction term (and with $\IR^d$ replaced by a torus) is considered 
by~\cite{lions/mas-gallic:2001} who use
it as a basis for a particle method for numerical solution of the porous medium equation.
Of course this does not quite fit into our framework, since in the notation of our 
population models this would necessitate $\gamma(x,m)=\smooth{\epsilon}m$ which is
not bounded. However, this can be overcome by an additional layer of approximation
(c.f.~our numerical experiments of Section~\ref{beyond linear diffusion})
and we do not allow this to detain us here. Existence and uniqueness of solutions 
to~(\ref{mollified equation}) can be obtained using the approach
of~\cite{lions/mas-gallic:2001}.

@@

\begin{corollary} \label{cor:wavefront}
    In addition to the assumptions of Corollary~\ref{cor:lineages_simple},
    suppose that the population process is described by a travelling wave with 
velocity $\wavespeed$,
    i.e., the population has density
    $\varphi(t, x) = w(x - t \wavespeed)$
    where $w$ solves
    \begin{align*}
        r \DG^* (\gamma w) + w F + \wavespeed \cdot \grad w = 
r\sigma^2\Delta (\gamma w) -\meanq \cdot\grad (\gamma w) +\wavespeed \cdot \grad w =0 .
    \end{align*}
    Then the semigroup $Q_s$
    of the motion of a lineages in frame that is moving at speed $\wavespeed$
    is time-homogeneous with generator
    \begin{align}
\label{Lgen}
        \Lgen f
        &=
        \sigma^2 r \gamma
        \left(
            \Delta f
            +
            2 \grad \log (\gamma w)
            \cdot \grad f
        \right)
        + (\wavespeed - r\gamma \meanq) \cdot \grad f .
    \end{align}
\end{corollary}

@@

Comparing %\eqref{PME} 
with \eqref{eqn:nonlocal_pde},
we see that to set up a limit in which the population density $\varphi$ follows the 
porous medium equation with logistic growth of~(\ref{PME}),
we need $r=1$,
$\gamma = \varphi$, and $F = 1 - \varphi$.
Consulting equation~\eqref{eqn:mu_defn},
this implies that $\mu_\theta = \max\big(0, (1 + 1/\theta) \varphi - 1/\theta\big)$.
In other words,
establishment is certain
and birth rates increase linearly with population density,
but to compensate, death rates increase slightly faster (also linearly),
as shown in Figure~\ref{fig:pme_waves}.
Alert readers will notice that
the condition from Assumptions~\ref{def:model_setup} 
that $\gamma(x,m)$ be uniformly bounded 
is violated.
This can be corrected by use of a cut-off, and in fact the downwards drift
provided by the logistic control of the population size prevents $m$ from 
getting too big. 
In practice in our simulations we
take discrete time steps of length $dt$ (with $dt$ suitably small),
and have each individual reproduce and die with probabilities, respectively,
\begin{align*}
    p_\text{birth}(m) = \left(1 - e^{- m dt}\right)
    \qquad 
    p_\text{death}(m) = \left(1 - e^{- (m (1+1/\theta) - 1/\theta) dt} \right) ,
\end{align*}
where $m$ is the local density at their location.
This makes $\gamma(x, m) = p_\text{birth}(m)/dt \approx m$
and
\begin{align*}
    F(x, m) = \theta(\gamma(x, m) - \mu(x, m)) / dt \approx 1 - m .
\end{align*}
Birth and death rates are equal at density $m = 1$,
corresponding to an unscaled density of $N$ individuals per unit area.

\begin{figure}
    \begin{center}
        \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
    \end{center}
    \caption{
        Simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        (implemented by choosing a small $dt$ taking reproduction probability per time step
        to be $1 - \exp(- dt m)$
        and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        values of $\theta$ are 1 and 100, respectively, and both values of $N$ are the same.
        \label{fig:pme_waves}
    }
\end{figure}

In one dimension, equation~(\ref{PME}) has an explicit travelling wave solution
\begin{align} \label{eqn:pme_wave}
    w^P(t, x)
    :=
    \left( 1 - e^{ \frac{1}{2} (x - x_0 - t) } \right)_+ .
\end{align}
Notice that the wave profile has a sharp boundary at $x = x_0 + t$.
There are also travelling wave solutions with $c>1$ \citep{gilding/kersner:2005},
which lack this property.
However, for initial conditions that decay sufficiently rapidly at infinity,
such as one might use in modelling a population invading new territory,
the solution converges to \eqref{eqn:pme_wave} \citep{kamin/rosenau:2004}.
In Figure~\ref{fig:pme_waves} we show simulations of the individual based model described above.
They display travelling wave solutions qualitatively similar to solutions of~(\ref{PME}),
with better agreement for smaller $\theta/N$
(but in both cases, $N$ is reasonably large).

@@

\paragraph{Porous Medium equation with logistic growth:}
Finally, consider equation~(\ref{PME}). Setting $x_0=0$ (for definiteness) and substituting
the form of $w^P$ from equation~\eqref{eqn:pme_wave}
into Corollary~\ref{cor:wavefront},
with $c=1$,
$\gamma(x, m) = m$,
$r(x,w) = 1$,
and $F(x, m) = (1 - m)$,
the generator of the diffusion governing the position of the ancestral lineage relative to the wavefront
is, for $x < 0$,
%with $f_x$ denoting the spatial derivative of $f$,
\begin{align*}
    \Lgen_3 f
    &=
        w^P
        \left(
        \partial_{xx} f
         +
         2 \frac{\partial_x((w^P)^2)}{(w^P)^2} \partial_xf
        \right)
        + \partial_xf \\
    &=
        \left(1 - e^{\frac{1}{2} x} \right)
        \partial_{xx}f
        -
        2 e^{\frac{1}{2} x} \partial_xf
        +
        \partial_xf .
\end{align*}
The speed measure corresponding to this diffusion has density
%again for $\xi < 0$,
\begin{align*}
    m(\xi)
    &\propto
        \frac{ 1 }{ 2 (1 - e^{\xi/2}) }
        \exp\left(
            \int_\eta^\xi \left\{
                1 - \frac{e^{x/2}}{1 - e^{x/2}}
            \right\} dx
        \right) \\
    &\propto
        e^\xi\left(1-e^{\xi/2}\right),
        \quad
        \text{for } \xi < 0 ,
\end{align*}
which is integrable and so when suitably normalised gives the unique stationary distribution.
Notice that even though we have the same reaction term as in the Fisher-KPP equation, with this 
form of nonlinear diffusion, at stationarity
the lineage will typically be significantly behind the front, 
suggesting a different genealogy.
